---
title: "Classification: LDA and QDA"
author: "Andrew Kerr"
format: 
  html:
    code-fold: true
    code-line-numbers: true
    code-tools: true
    self-contained: true
editor: visual
execute:
  message: false
---

## Setup

Declare your libraries:

```{r}
#| label: libraries-r
#| include: false
library(tidyverse)
library(tidymodels)
library(glmnet)
library(discrim)
```

# LDA

```{r}
ins <- read_csv("https://www.dropbox.com/s/bocjjyo1ehr5auz/insurance.csv?dl=1")

ins <- ins %>%
  mutate(
    smoker = factor(smoker)
  ) %>%
  drop_na(smoker)
```

```{r}
lda_mod <- discrim_linear() %>%
  set_engine("MASS") %>%
  set_mode("classification")
```

```{r}
lda_fit_1 <- lda_mod %>%
  fit(smoker ~ charges, data = ins)

lda_fit_1$fit %>% summary()
```

```{r}
preds <- predict(lda_fit_1, new_data = ins)

ins <- ins %>%
  mutate(
    pred_smoker = preds$.pred_class
  )

ins %>%
  accuracy(truth = smoker,
           estimate = pred_smoker)
```

```{r}
lda_fit_2 <- lda_mod %>%
  fit(smoker ~ charges + age, data = ins)

lda_fit_2
```

```{r}
lda_fit_2$fit$scaling
```

```{r}
ins %>%
  ggplot(mapping = aes(x = charges, y = age, color = smoker)) +
  geom_point()
```

```{r}
lda_fit_2

coefficients <- lda_fit_2$fit$scaling

my_slope = coefficients[1] / 
              (-1 *coefficients[2])
```

```{r}
ins %>%
  ggplot(mapping = aes(x = charges, y = age, color = smoker)) +
  geom_point() +
  geom_abline(mapping = aes(slope = my_slope, intercept = 0), 
              lwd = 1.5) +
  scale_x_continuous(labels = label_dollar()) +
  labs(x = "Medical Charges", 
       y = "Age", 
       color = "Smoking Status")
```

#### Your turn:

Use cross-validation to find the best LDA model to predict smoker status.

```{r}
ins_cvs <- vfold_cv(ins, v = 10)

################################################################################

recipe_1 <- recipe(smoker ~ ., data = ins) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_nominal_predictors())

recipe_2 <- recipe(smoker ~ charges + bmi + age, data = ins) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_nominal_predictors())

recipe_3 <- recipe(smoker ~ charges + bmi + age + sex, data = ins) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_normalize(all_nominal_predictors())

################################################################################

workflow_1 <- workflow() %>%
  add_model(lda_mod) %>%
  add_recipe(recipe_1)

workflow_2 <- workflow() %>%
  add_model(lda_mod) %>%
  add_recipe(recipe_2)

workflow_3 <- workflow() %>%
  add_model(lda_mod) %>%
  add_recipe(recipe_3)

################################################################################

fit_1 <- workflow_1 %>% fit_resamples(ins_cvs)
fit_2 <- workflow_2 %>% fit_resamples(ins_cvs)
fit_3 <- workflow_3 %>% fit_resamples(ins_cvs)

################################################################################

fit_1 %>% collect_metrics()
fit_2 %>% collect_metrics()
fit_3 %>% collect_metrics()
```

How does it compare to the Logistic Regression and KNN approaches?

Logistic regression has a higher accuracy, but Logistic Regression and LDA have about the same ROC AUC with KNN having a slightly higher value.

# Quadratic Discriminant Analysis

#### Code:

```{r}
qda_mod <- discrim_regularized(frac_common_cov = 0) %>% 
             set_engine("klaR") %>% 
             set_mode("classification")
```

#### Your turn:

Use Cross-Validation to find the best QDA model to predict smoker status.

```{r}

workflow_1 <- workflow() %>%
  add_model(qda_mod) %>%
  add_recipe(recipe_1)

workflow_2 <- workflow() %>%
  add_model(qda_mod) %>%
  add_recipe(recipe_2)

workflow_3 <- workflow() %>%
  add_model(qda_mod) %>%
  add_recipe(recipe_3)

################################################################################

fit_1 <- workflow_1 %>% fit_resamples(ins_cvs)
fit_2 <- workflow_2 %>% fit_resamples(ins_cvs)
fit_3 <- workflow_3 %>% fit_resamples(ins_cvs)

################################################################################

fit_1 %>% collect_metrics()
fit_2 %>% collect_metrics()
fit_3 %>% collect_metrics()
```

How does it compare to the LDA, Logistic Regression, and KNN approaches?

This approach has a slightly worse accuracy, but a ROC AUC on par LDA and Logistic Regression.
