---
title: "Final Project"
author: "Lily Cook, Andrew Kerr, Daniel Erro"
format: html
editor: visual
embed-resources: true
---

```{r}
#| label: libraries
#| message: false

library(tidyverse)
library(here)
library(tidymodels)
library(tidyclust)
library(magrittr)
library(readxl)
library(janitor)
```

# College Rankings

## Data Read-in

```{r}
#| label: rankings-read-in
#| message: false

rank <- read_xlsx(here('project', 
                       'data', 
                       'US-News-National-University-Rankings-Top-150-Through-2025.xlsx')) %>%
  clean_names()
```

## Data Cleaning

```{r}
#| label: rankings-cleaning

rank %>%
  summarise(across(everything(), ~ round(sum(is.na(.)) / nrow(rank), 2))) %>%
  pivot_longer(cols = everything(), 
               names_to = 'Year', 
               values_to = 'Proportion_Missing') %>%
  arrange(-Proportion_Missing)

rank %>%
  select(-university_name, -state, -ipeds) %>%
  summarise(across(everything(), ~ max(., na.rm = T))) %>%
  pivot_longer(cols = everything(), 
               names_to = 'year', 
               values_to = 'max_rank') %>%
  mutate(year = as.numeric(gsub("x", "", year))) %>%
  ggplot() +
  geom_col(aes(x = year, y = max_rank), fill = 'cornflowerblue') +
  theme_bw() +
  labs(x = "Year",
       y = "Maximum Ranking",
       title = "Maximum Rankning by Year")


```

# College Scorecard

## Data Read-in

```{r}
#| label: scorecard-read-in
#| message: false

card <- read_csv(here('project', 'data', 'Most-Recent-Cohorts-Institution.csv'))
```

NOTES: - What year variable values are from: CollegeScorecardDataDirectory \> Most_Recent_Inst_Cohort_Map - What each variable means: Technical Documentation: College Scorecard Institution-Level Data - They generalized the names, so search for just the first part of each variable

## Data Cleaning

```{r}
#| label: card-cleaning

card %>%
  summarise(across(everything(), ~ round(sum(is.na(.)) / nrow(card), 2))) %>%
  pivot_longer(cols = everything(), 
               names_to = 'Column', 
               values_to = 'Proportion_Missing') %>%
  arrange(-Proportion_Missing)

joined_data <- left_join(rank, card, by = join_by(university_name == INSTNM))

```

```{r}
#| label: check-join

r_lst <- rank %>%
  select(ipeds) %>%
  unique()

c_lst <- card %>%
  select(UNITID) %>%
  unique()

nrow(r_lst)
nrow(c_lst)

nrow(inner_join(r_lst, c_lst, by = join_by(ipeds == UNITID)))
anti_join(r_lst, c_lst, by = join_by(ipeds == UNITID))

# Drop this observation because it is a branch of NYU that only existed separately for 3 years, and NYU is already in the dataset
rank %>%
  filter(ipeds == 194541)
```

# Functions

```{r}
#| label: test-k-function

fit_kmeans <- function(data, recipe, k, counts = F) {

  km_spec <- k_means(num_clusters = k)
  
  km_wflow <- workflow() %>%
    add_recipe(recipe) %>%
    add_model(km_spec)
  
  km_fitted <- km_wflow %>% fit(data)

  engine_fit <- km_fitted %>% 
    extract_fit_engine()
  
  df_results <- data.frame(
    k = k,
    Cluster = paste0("Cluster_", seq(1:k)),
    Withiness = engine_fit$withinss,
    Betweenss = engine_fit$betweenss,
    Total_Withiness = engine_fit$tot.withinss
  ) %>% pivot_wider(names_from = Cluster, values_from = Withiness)
  
  if(counts) {
    
    df_counts <- data.frame(
      k = k,
      extract_cluster_assignment(km_fitted) %>%
        group_by(.cluster) %>%
        count()
    ) %>% pivot_wider(names_from = .cluster, values_from = n)
    
    return(list(results = df_results, counts = df_counts))
    
  }
    
  return(results = df_results)
  
}
```

# KMeans

```{r}
#| label: kmeans

card_recipe_pca <- recipe(~., data = card %>% select(where(is.numeric))) %>%
  # update_role(INSTNM, new_role = "id") %>%
  # step_rm(UNITID, OPEID, OPEID6, CITY, STABBR, ZIP, ACCREDAGENCY, INSTURL, NPCURL, MAIN, NUMBRANCH) %>%
  step_select(where(~ !all(is.na(.)))) %>%
  step_impute_mean(all_numeric()) %>%
  step_zv(all_predictors()) %>%
  # step_unknown(all_nominal()) %>%
  # step_dummy(all_nominal()) %>%
  step_center(all_numeric()) %>%
  step_scale(all_numeric()) %>%
  step_pca(all_numeric(), threshold = 0.8)

test <- card_recipe_pca %$%
  steps %>% 
  pluck(7) %$%
  res

test$rotation[,'PC1']

k_lst <- c(5, 10)

kmeans_results <- list_rbind(map(k_lst, ~fit_kmeans(card %>% select(where(is.numeric)), card_recipe_pca, .x)))

kmeans_results
```
