---
title: "Assignment 2: Cross-Validation and K-Nearest-Neighbors"
author: "Andrew Kerr"
format: 
  html:
    code-fold: true
    code-line-numbers: true
    code-tools: true
    embed-resources: true
editor: source
---

```{r}
#| label: packages
#| message: false
#| warning: false

set.seed(3849)

library(tidyverse)
library(tidymodels)
library(kableExtra)
```

# Instructions

Your document should also be clearly organized, so that it is easy for a reader to find your answers to each question. If I have a difficult time locating or reading your answer, you risk it being returned with a "revision requested".

# The Data

```{r}
#| label: load-data
#| message: false
whr_clean <- read_csv(here::here('data', 'whr_clean.csv'))

whr_clean <- whr_clean %>%
  mutate(
    regional_indicator = as_factor(regional_indicator)
  )

whr_clean %>% summary() %>% kable()
```

# Happiness Scores

## Part 1: Happiness Over Time

1.  Is the happiness in the world changing linearly over time? Fit a simple linear model and interpret the results to address this question.

```{r}
#| label: Q1

lin_reg <- linear_reg() %>%
  set_engine("lm") %>%
  set_mode("regression")

lr_m1 <- lin_reg %>%
  fit(happiness_score ~ year, data = whr_clean)

whr_clean <- whr_clean %>%
  mutate(
    prediction = predict(lr_m1, new_data = whr_clean)$.pred,
    residuals = happiness_score - prediction
    )

glance(lr_m1) %>% kable()
tidy(lr_m1) %>% kable()

whr_clean %>%
  ggplot(aes(x = happiness_score, y = residuals)) +
    geom_point() +
    theme_bw()
```

The happiness in the world is not changing linearly over time since we see a low
$R^2$ value of approx. 0, and a large p-value for the year coefficient. 
Additionally, the residual plot shows a linear pattern, meaning we are using 
the wrong model.

2.  Was the happiness score approximately the same in all the years? Convert `year` to a factor variable, and fit a simple linear model to address this question.

```{r}
#| label: Q2

whr_clean <- whr_clean %>%
  mutate(year = as_factor(year))

lr_m2 <- lin_reg %>%
  fit(happiness_score ~ year, data = whr_clean)

tidy(lr_m2) %>% kable()
```
 No, the happiness score was not approx. the same in all the years. We see 
 highly significant p-values for all years compared to the baseline year of 
 2005, with 2005 being on average around 1 unit of happiness score greater than
 all the other years.

## Part 2: Happiness Equation

3.  How is each of the six measures of life quality weighted in calculating this score? Fit a model to estimate the weights, and interpret the coefficients.

```{r}
#| label: Q3

lr_m2 <- lin_reg %>%
  fit(happiness_score ~ year, data = whr_clean)
```

4.  Which measures of life quality does the WHR consider to be most important to a country's happiness?

# Predicting Life Expectancy -- Exploring Linear Models

## Part 1: Summary of Approach

6.  Write a short description (bullet points are fine) of your process in narrowing down your model. How did you approach this problem, without spending hours upon hours fitting and cross-validating zillions of models?

## Part 2: Three Candidates

8.  Choose the three best candidate models among those you tried.

9.  Supply your code and results for comparing these models, and discuss how you decided which *one* model was the best one.

## Part 3: Final Model

10. Summarize the results from your final model. Don't forget to fit your final model on the **full** dataset, even if you used *test / training* data or *cross-validation* during the model selection process.

11. Include a plot of the residuals and discussion of what you see, and interpretations of the coefficients and metrics.

# Predicting Life Expectancy -- K-Nearest-Neighbors

## Part 1: Tuning K

12. For **each** of your top three candidate models from Q8, find the best choice of **K**. Show all your work, and provide a brief summary at the end (e.g., "For Model 1, we choose a K of \[something\] because \[reasons\]. For Model 2, ...")

## Part 2: Best Model

13. Fit and report your single best model from Q9.

You should include:

-   An argument for your choice of K, including a plot.

-   A plot of the residuals

# Predicting on New Data

14. Use your **one** best *least-squares regression* model (Q9) to predict the life expectancy of all countries.

15. Use your **one** best *KNN* model (Q13) to predict the life expectancy of all countries.

16. Which model did a better job predicting the true values in the new data?

# Discussion Questions

## Parametric and Non-Parametric

17. Make an argument for why a **parametric** method, like least-squares regression, might be preferred in this task.

18. Then make an argument for why a **non-parametric** method, like K-Nearest-Neighbors, might be preferred.

## Interpretation and Prediction

19. If your only goal were **interpretation**, which of the candidate models (from *any* section of the assignment) would you have chosen? Why?

20. If your only goal were **prediction** on future data, which of the candidate models would you have chosen? Why?

## Standardization

21. Consider your final best least-squares regression model from Q9. Suppose we fit this model again, but this time, we normalize **all** of the quantitative variables. Will anything change? Give a (conceptual) argument for why this is true.

## Quantitative or Categorical?

22. Suppose we add the predictor `year` to our final model as a **categorical variable** and fit the model on all the data. Then, we use this new model to predict on the 2020 data. What is wrong with this plan?
